---
title: "tRAT3 Appeals"
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

### #33
For #33, c is the correct answer.
The code for c:
```{r}
p_33 <- read_csv("a,b,c\n1,2,4\n1,2,3", col_names = TRUE)
p_33
is_tibble(p_33)
```
We see that it is indeed a 2x3 tibble.

### #37
For 37, c is correct. A, b, and d are the defining characteristics of tidy data. E is arguably correct as well, as tibbles are not mandatory for tidy data, though they do make it easier to tidy up data, and to use tidy data. We argue that c is correct as it is the most wrong: in tidy data, each row contains one observation, which may (and almost always will) consist of many variables. 

### #38
For #38, the answer is b because gender should be gathered into one column. Gender is one observation and thus in optimal tidy data would have its own column. A and c are wrong as they are characteristics of tidy data, d is wrong as it prints that it is indeed a tibble. E is wrong, as the missing values are present in the table and therefore are explicit, not implicit.

### #39
For #39, b is correct.

Building the dataframe and running the code from b, we get a tidy tibble which allows us to use the first and last names individually.
```{r}
tb <- tibble(
  person = c("Nelson, Jacqi","Nelson, Jacqi","Nelson, Jacqi","Winfield, Ron", "Winfield, Ron", "Winfield, Ron", "Zhao, Julia"),
  treatment = c(1.00,2.00,3.00,1.00,2.00,3.00,1.00),
  response = c(7.00, 10.0, 9.00, 4.00, 5.00, 8.00, 4.5)
)
tb
separate(tb,person, into=c("last","first"),sep = ",")
```



