---
title: 'Lab 6: Airport Data Science'
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
COflights <- read_csv("COflights.csv")
```

### What Factors Lead to On-time Arrivals in Denver?

**Why this question is important:**

**Why the answer is relevant:**

### Team Plot and Findings

```{r}
#Team Plot
```

### Recommendations

How to reduce flight delays/why implement this change

### Individual Sections

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights



#### Individual Contributions

**Name:**

**Name:**

**Name:**

**Name:**
