---
title: 'Lab 6: Airport Data Science'
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
COflights <- read_csv("COflights.csv")
```

### What Factors Lead to On-time Arrivals in Denver?

**Why this question is important:**
This question is important when trying to study which facotrs affect a delated flight. Undertsnading these factors allows airports or carriers to adapt and make changes based on the factors we found..... needs more

**Why the answer is relevant:**
This question is relavant for the DIA managers who want to increase the amount of on-time arrivals and which factors lead to delayed flights. These factors could include either the weather or the carriers fault, but the managers want to know how to minimize these delayed arrivals and increase the amount of on-time flights. 

### Team Plot and Findings

```{r}
#Team Plot
```

### Recommendations

How to reduce flight delays/why implement this change

### Individual Sections

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights

#### Name's Plot

```{r}

```

Identify 2 factors from the data that lead to more or less delayed flights



#### Individual Contributions

**Name:**

**Name:**

**Name:**

**Name:**
