---
title: "Lab 4: Smoking and Low Birth Weight?"
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"             
date: "9/20/2019"
output: html_document
---
```{r setup,include=FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
```

### The Main Questions

**The Question:** Does Smoking Lead to Babies Born Prematurely and with Low Birth Weight?

#### Premature births

**Why the question is important:** 

**Why the answer is relevant:**

#### Birth weight

**Why the question is important:** 

**Why the answer is relevant:**

#### Analysis of premature births

```{r Team Plot1,echo=TRUE, warning=FALSE}

```

**Surgeon General's Assertion:** Mothers who smoke have increased rates of premature delivery (before 270 days).

**The answer:** 

#### Analysis of birth weights at gestational ages

```{r Team Plot2,echo=TRUE, warning=FALSE}
bool_smoke <- as_labeller(c(`0` = "Non Smokers", `1` = "Smokers"))
ggplot(na.omit(babies))+
  stat_density_2d(aes(x=gestation,y=bwtoz,fill=..level..),geom = "polygon")+
  #xlim(200,375)+
  geom_vline(aes(xintercept=270),color="red")+
  facet_wrap(~smoke,labeller=bool_smoke)+
  xlab("Gestation (days)")+
  ylab("Birth weight (ounces)")+
  scale_fill_continuous(name="Proportion")
```

**Surgeon General's Assertion:** The newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born).

**The answer:** 

### Recommendations 

Summarize answers to surgeon general's assertions above

Summarize individual sections

### Preliminary Question

**The Question:** Which variables are similar between smokers and nonsmokers? Which are different?

```{r Team Plot3,echo=TRUE, warning=FALSE}
plot3 <- ggplot(na.omit(babies))
#plot3+geom_bar(aes(med))+facet_wrap(~smoke) less higher education among smokers also plot3+geom_bar(aes(ded))+facet_wrap(~smoke)
# plot3+geom_density(aes(mage))+facet_wrap(~smoke) maybe mother age
#plot3+geom_bar(aes(parity))+facet_wrap(~smoke) fewer smokers with low num of pregnancies
# plot3+geom_density(aes(mpregwt))+facet_wrap(~smoke) maybe something
#plot3+geom_density(aes(dwt))+facet_wrap(~smoke) similar to mothers
```


### Individual Findings and Plots

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**
 
### Summary of Individual Contributions
<!--Note what non default aesthetics and layers used--> 
**Brian Teklits:** 

**Andrew:** 

**Clint:** 

**Charles:** 
