---
title: "Lab 4: Smoking and Low Birth Weight?"
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"             
date: "9/20/2019"
output: html_document
---
```{r setup,include=FALSE}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)
```

### The Main Questions

**The Question:** Does Smoking Lead to Babies Born Prematurely and with Low Birth Weight?

#### Premature births

**Why the question is important:** 

  It is important to ask such questions in order to fully understand the affects of smoking to babies. By researching and analyzing data we can notice trends that will allows us to clearly see the results of smoking. We want to analyze this data in order to warn and inform women of the consequences. If smoking truly leads to premature birth than women should be aware of their choices and morale obligations. 

**Why the answer is relevant:**

  We focus on this question in order to change the future of pregnant smoking. By analyzing trends and examining these questions, we can inform women about the dangers of smoking while pregnant and hopefully prevent any unwarrented problems with new borns.
  
#### Birth weight

**Why the question is important:** 

**Why the answer is relevant:**

#### Analysis of premature births

```{r Team Plot1,echo=TRUE, warning=FALSE}

```

**Surgeon General's Assertion:** Mothers who smoke have increased rates of premature delivery (before 270 days).

**The answer:** 

#### Analysis of birth weights at gestational ages

```{r Team Plot2,echo=TRUE, warning=FALSE}
bool_smoke <- as_labeller(c(`0` = "Non Smokers", `1` = "Smokers"))
ggplot(na.omit(babies))+
  stat_density_2d(aes(x=gestation,y=bwtoz,fill=..level..),geom = "polygon")+
  #xlim(200,375)+
  geom_vline(aes(xintercept=270),color="red")+
  facet_wrap(~smoke,labeller=bool_smoke)+
  xlab("Gestation (days)")+
  ylab("Birth weight (ounces)")+
  scale_fill_continuous(name="Proportion")
```

**Surgeon General's Assertion:** The newborns of mothers who smoke have smaller birth weights at every gestational age (number of days into pregnancy when child is born).

**The answer:** 

### Recommendations 

Summarize answers to surgeon general's assertions above

Summarize individual sections

### Preliminary Question

**The Question:** Which variables are similar between smokers and nonsmokers? Which are different?

```{r Team Plot3,echo=TRUE, warning=FALSE}
plot3 <- ggplot(na.omit(babies))
#plot3+geom_bar(aes(med))+facet_wrap(~smoke) less higher education among smokers also plot3+geom_bar(aes(ded))+facet_wrap(~smoke)
# plot3+geom_density(aes(mage))+facet_wrap(~smoke) maybe mother age
#plot3+geom_bar(aes(parity))+facet_wrap(~smoke) fewer smokers with low num of pregnancies
# plot3+geom_density(aes(mpregwt))+facet_wrap(~smoke) maybe something
#plot3+geom_density(aes(dwt))+facet_wrap(~smoke) similar to mothers
```


### Individual Findings and Plots

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**

#### Individual Plot

```{r  echo=TRUE, warning=FALSE}


```

**Individual Findings:**
 
### Summary of Individual Contributions
<!--Note what non default aesthetics and layers used--> 
**Brian Teklits:** 

**Andrew:** 

**Clint:** 

**Charles:** 
