---
title: 'Lab 15: Final Project'
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"
date: "12/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,message=FALSE)
library(modelr)
library(tidyverse)
library(maps)
library(lubridate)
# install.packages("openintro")
library(openintro)
# Data from: https://www.kaggle.com/sobhanmoosavi/us-accidents
# LARGE: do not upload to github
accidents <- read_csv("US_Accidents_May19.csv")

```

## Main Question

**Domain Expert and Dataset:**

The domain expert is the NHTSA, the National Highway Traffic Safety Administration.

The dataset we are using can be found [here.](https://www.kaggle.com/sobhanmoosavi/us-accidents) It contains 2.25 million records of accidents from 49 different states, starting in February 2016 until March 2019. It includes data on the severity of the accident, where and when the accident occurred, the weather at the time of the accident, and various information about the nearby surroundings. 

**Main Question & it's Importance:**
Lead with your main overall question, why it’s important (and why those reasons are important), 

**Answer:**
your answer/conclusion, 
**Recommendation:**
```{r}

```

recommendation, and a plot communicating your conclusion or basis for your recommendation. 

## Individual Sections

Each teammate must apply “recently learned” programming tools to answer a subquestion that will help your team answer the overall question. The subquestions must be different from the overall team question. Each subquestion should get its own subsection in the team report and should follow the format for statistical analyses that we have learned in class.

What is the question? Why is it important/interesting?

What new tools did you use?

What is the answer to the question? How does this relate to the overall question?

What are the ethical implications of your findings, conclusions, or recommendations?

### Andrew

### Brian

### Charles

```{r}
cd_data <- accidents %>%
  mutate(a_start = ymd_hms(Start_Time)) %>%
  mutate(a_end = ymd_hms(End_Time)) %>%
  mutate(accident_time = difftime(a_end, a_start, units = "mins")) %>%
  filter(!is.na(`Humidity(%)`), !is.na(accident_time)) %>%
  filter(accident_time >= 0 & accident_time <= 4320) %>%
  select(a_start, a_end, accident_time, `Humidity(%)`)
```

```{r}
humid_time_model <- lm(`Humidity(%)` ~ cd_data$accident_time, data = cd_data)

humid_preds <- cd_data %>%
  data_grid(`Humidity(%)` = seq_range(`Humidity(%)`, 2178823)) %>%
  add_predictions(humid_time_model,'predictions') %>%
  add_residuals(humid_time_model, 'residuals')
```

```{r}
ggplot() +
  geom_point(data = cd_data, aes(`Humidity(%)`,accident_time))
```

### Clint

## Reflections on Lab 2

As a team, restate your team’s goal(s) and reflect on how well you achieved it. If you could travel back in time, which is one thing you would tell your team to keep doing? Stop doing? Start doing? Discuss this as a team and add it as a section to your Final Project.

### Individual Reflections

Individually, write a paragraph reflecting on: how have your six-month or 5-year goals changed? What did you learn/accomplish in this course? If you could give yourself advice at the beginning of the semester, what would you tell yourself to keep doing, stop doing, and start doing?

**Andrew:**

**Brian:**

**Charles:** 

**Clint:**

### Individual Contributions

**Andrew:**

**Brian:**

**Charles:** 

**Clint:**