---
title: "Lab 3: Plotting Health Data"
author: "Brian Teklits, Charles Doremieux, Andrew MacLean"               #add your names
date: "9/10/2019"
output: html_document
---
```{r include=FALSE}
library(tidyverse)
library(scales)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

### The Overall Question

**The Question:** How does the amount of weight a person wants to lose relate to their overall current health? 

**Why the question is important:** For these two factors to be related would suggest respondents are aware their health would improve if they reached their desired weight, but are unable to enact change. Exploring this question could suggest which other data vectors relate to stagnation in health. (e.g. Healthplan/gender/smoking/age/exercise/height)

**Why the answer is relevant:** By exploring the topic of awareness of one's health we gain insight into the habits of healthy individuals and how they differ from those in poor health. In-depth analysis opens the door to recommendations for possible change. What factors (healthplan/smoking/exercise) or groups (Age/gender) could be targeted? 

### Team Plot and Conclusions


```{r Team Plot,echo=FALSE, warning=FALSE}
ggplot(cdc)+
  geom_density(aes(x=(wtdesire-weight),y=..density..,fill=genhlth),position="stack")+
  xlim(-75,25)+
  xlab("Desired Change in Weight (lbs)")+
  ylab("Percentage of Respondents")+
  scale_fill_discrete(name="General Health")+
  scale_y_continuous(labels=scales::percent) #from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
```

**The answer:** 
Just over 25% of respondents are happy at their current weight. As we move away from this value, we see that respondents in the categories of poor, fair, and good makeup the bulk of the proportion. These people represent those who would like to make a meaningful change in their overall health, but have been unable to do so thus far. This group should be targeted in policy decisions to maximize their effect.

### Recommendations 

Exercise good
Healthplan good
smoke bad?

age or gender to target

### Individual Findings and Plots

#### Brian

```{r Brian, echo=FALSE, warning=FALSE}
health_plan_label <- as_labeller(c(`0` = "No Health Plan", `1` = "Has a Health Plan")) #from https://ggplot2.tidyverse.org/reference/as_labeller.html
ggplot(cdc)+
  geom_boxplot(aes(x=genhlth,y=(wtdesire-weight)),outlier.alpha = .3,outlier.shape = 21,outlier.color = "red")+
  ylim(-175,125)+
  coord_flip()+
  facet_wrap(~hlthplan,labeller = health_plan_label)+
  ylab("Desired Change in Weight (lbs)")+
  xlab("General Health")

```

**Individual Findings:**
Interestingly, we find that people who have a health plan generally want to change their weight more than those who don't. There are both more outliers, and a wider range in the confidence intervals for the health plan data set, compared to the no health plan data. 

This suggests that people who are actively engaged in monitoring their health are better informed on what overall changes they need to make. Engaging the group with no health plan with low cost alternatives, such as health fairs, could help to inform them of what beneficial changes they can make. 

Additionally, for those people with healthcare, encouraging doctors to refer patients to dieticians, when health plans allow, could help them to plot a path to their desired weight and health goals. 

#### Andrew's Plot

```{r Name2 Plot, echo=FALSE, warning=FALSE}
# Andrew's Plot
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
ggplot(data = cdc, mapping = aes(x = weight, y = wtdesire-weight))+
  geom_smooth(method = 'lm') +
  geom_point(mapping = aes(color = gender))+
  ylim(-200,100)+
  facet_grid(genhlth ~ exerany)

```

**Individual Findings:**
This graph represents that there is a tendency that people who are of heavier weight and work out on the regular are usually unhappy with their current weight. The trend of the best fit line explains that people of poor health are more likely to be unhappy with their current weight and desire to lose weight. The majority of people no matter their general health usually prefer to gain a few pounds or stay at their current weight. Despite the status of their health, it seems the majority of people who workout only want minimal changes to their weight.

#### Name Here

```{r Name3 Plot, echo=FALSE, warning=FALSE}
# Individual Plot
```

**Individual Findings:**
How does this support overall conclusion

#### Name Here

```{r Name4 Plot, echo=FALSE, warning=FALSE}
# Individual Plot
```

**Individual Findings:**
How does this support overall conclusion
 
### Summary of Individual Contributions

**Name:**

