---
title: "Lab 3: Plotting Health Data"
author: "Brian Teklits, Charles Doremieux, Andrew MacLean"               #add your names
date: "9/10/2019"
output: html_document
---
```{r include=FALSE}
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
```

### The Overall Question

**The Question:** How does the amount of weight a person wants to lose relate to their overall current health? 

**Why the question is important:** If they are related, it suggests respondents are aware their health would improve if they reached their desired weight, but are unable to enact change. Why? Which other data vectors relate to this? Healthplan/gender/smoking/age/exercise/height?

**Why the answer is relevant/important:** We can gain insights into how the habits of healthier individuals differ from those in poorer health, and use this to make recommendations on possible change. What factors (healthplan/smoking/exercise) or groups (Age/gender) could be targeted? 

### Team Plot and Conclusions


```{r Team Plot,echo=FALSE, warning=FALSE}
ggplot(cdc)+
  geom_density(aes(x=(weight-wtdesire),y=..density..,fill=genhlth),position="stack")+
  #ylim(0,20)+
  xlim(-25,75)+
  xlab("Desired Change in Weight(lbs)")+
  ylab("Proportion of Respondents")
  #facet_wrap(~exerany)
```


**The answer:**

### Recommendations 

Recommendations based on the answer

### Individual Findings and Plots

#### Name Here

```{r Name1 Plot, echo=FALSE, warning=FALSE}
# Individual Plot
```

**Individual Findings:**
How does this support overall conclusion

#### Andrew's Plot

```{r Name2 Plot, echo=FALSE, warning=FALSE}
# Andrew's Plot
library(tidyverse)
source("http://www.openintro.org/stat/data/cdc.R")
cdc <- as_tibble(cdc)
ggplot(data = cdc, mapping = aes(x = weight, y = wtdesire-weight))+
  geom_smooth(method = 'lm') +
  geom_point(mapping = aes(color = gender))+
  ylim(-200,100)+
  facet_grid(genhlth ~ exerany)

```

**Individual Findings:**
This graph represents that there is a tendency that people who are of heavier weight and work out on the regular are usually unhappy with their current weight. The trend of the best fit line explains that people of poor health are more likely to be unhappy with their current weight and desire to lose weight. The majority of people no matter their general health usually prefer to gain a few pounds or stay at their current weight. Despite the status of their health, it seems the majority of people who workout only want minimal changes to their weight.

#### Name Here

```{r Name3 Plot, echo=FALSE, warning=FALSE}
# Individual Plot
```

**Individual Findings:**
How does this support overall conclusion

#### Name Here

```{r Name4 Plot, echo=FALSE, warning=FALSE}
# Individual Plot
```

**Individual Findings:**
How does this support overall conclusion
 
### Summary of Individual Contributions

**Name:**

