---
title: "Lab 8: Data Science to Solve Problems and Answer Questions, part 1"
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"
date: "10/18/2019"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
```

### Dataset and Questions

**Who is hiring us?:** A candy company seeking to develop new products.

**Overall question and it's importance:** What candies are the most universally liked? 

Understanding which candies are the most popular informs us of what qualities to keep in mind when designing a new candy.

**Why the answer matters:** There are a variety of conclusions which can be drawn depending on what the company wants their new product to target. Universally liked candies offer a safe bet, universally disliked offer some qualities to avoid, while those in between could represent somewhat effective ideas which can be improved upon.

**Description of Dataset:** The dataset can be found [here.](https://www.scq.ubc.ca/wp-content/uploads/2017/10/candyhierarchy2017.csv) It contains information from a survey, with user data on gender, age, and country/state. Users rate candy on a scale of 3: DESPAIR, MEH, JOY. Essentially, these categories translate to dislike, indifferent, or like. 

There are also many other columns describing a variety of things added onto the survey, many of which are unclear what they describe. The candy data corresponds to 104 columns, each describing a different candy and the ratings associated to it from the 2460 respondents to the survey. Mixed in with the candy ratings are a variety of gag questions.

### Individual Questions

Some ideas: 
genders, age, country/state
popular/unpopular candy by demographic

**Andrew's Question:**

**Why this question is interesting:**

**Brian's Question:**

**Why this question is interesting:**

**Charles' Question:**

**Why this question is interesting:**

**Clint's Question:**

**Why this question is interesting:**

### East Africa Lake Data Visualizations

#### Andrew's Plot
```{r}
# CRU and MAM rainfall
```

#### Brian's Plot
```{r}
# VOI rainfall
```

#### Charles' Plot
```{r}
# Challa
```

#### Clint's Plot
```{r}
# Malawi
```