---
title: "Lab 5"
author: "Brian Teklits, Charles Doremieux, Andrew MacLean, Clint LaBattaglia"
date: "9/25/2019"
output: html_document
---

```{r setup,include=FALSE}
library(tidyverse)
milk <- read.csv('state_milk_production.csv')
```

All the writeup starter code:
```{r}
milk <- milk %>%
  mutate(milk_million = milk_produced/1000000)
```
```{r}
milk1970 <- milk %>%
  filter(year == 1970)
```

```{r}
ggplot(data = milk1970, aes(x = milk_million)) +
  geom_histogram(bins = 15) + 
  ggtitle('Histogram of Milk Produced in 1970 by State')+
  xlab("millions of gallons of milk")


ggplot(data = milk1970, aes(x = state, y = milk_million)) +
  geom_col() + 
  ggtitle('Histogram of Milk Produced in 1970 by State')+
  xlab("State")+
  ylab("millions of gallons of milk")+
  theme(axis.text.x = element_text(angle=60, hjust=1))+
  geom_text(aes(label = milk_million), vjust = -0.5)
  
```
```{r}
ggplot(data = milk1970, aes(x = milk_million)) +
  geom_density() + 
  ggtitle('Distribution estimate of milk produced in 1970 by state')+
  xlab("millions of gallons of milk")
```
```{r}
boxMilk <- milk %>%
  filter(year %in% c('1970','1990','2010'))
```

```{r}
ggplot(data = boxMilk, aes(x = as.factor(year), y = milk_million, fill = as.factor(year))) +
  geom_boxplot() + 
  ggtitle('Pounds of milk produced at the state-level') + 
  scale_fill_discrete(name = 'Year') + 
  xlab('Year') +
  ylab('Milk Produced (Millions lb)')
```

```{r}
milksub <- milk %>%
  filter(state %in% c('New York','Wyoming','Maryland','Maine','Florida')) %>%
  select(state, year, milk_million)
```
```{r}
ggplot(data = milksub, aes(x = year, y = milk_million, color = state)) +
  geom_point() + 
  ggtitle('Pounds of milk over time by state') +
  xlab('Year') +
  ylab('Milk Produced (Million lb)')
```
```{r}
cor_sum <- milksub %>%
  group_by(state) %>%
  summarise(Correlation = cor(year,milk_million)) %>%
  as_tibble()
cor_sum
```
Lab Starts here, remove above before submitting

### Team Plot
```{r}
# At least 5 states we've lived in: Colorado, California, New York
# For these states, plot a yearly time series for the milk production (in millions of pounds). On this plot, display a line showing the average milk produced each year for all states
```

**What does this plot convey:**

#### Reflections

**What was most helpful in learning about this data:**

**Who could benefit the most from this analysis? Who could be harmed by it?:**

### Individual Sections

#### Name's Plot

```{r}
# Yearly milk production for birth year
```

```{r}
# summarize data for birth year
 
```

report using summarise: avg milk produced per state, median milk produced by each state, state that produced the most milk, state that produced the least milk

#### Name's Plot

```{r}
# Yearly milk production for birth year
```

```{r}
# summarize data for birth year
 
```

report using summarise: avg milk produced per state, median milk produced by each state, state that produced the most milk, state that produced the least milk

#### Name's Plot

```{r}
# Yearly milk production for birth year
```

```{r}
# summarize data for birth year
 
```

report using summarise: avg milk produced per state, median milk produced by each state, state that produced the most milk, state that produced the least milk

#### Name's Plot

```{r}
# Yearly milk production for birth year
```

```{r}
# summarize data for birth year
 
```

report using summarise: avg milk produced per state, median milk produced by each state, state that produced the most milk, state that produced the least milk